Competing Bandits: \\ The Perils of Exploration under Competition

### 2018-10-18 paper/content/barriers.tex
somewhat major revision

\section{Data and Reputation as Barriers to Entry}\label{section:6}

Under temporary monopoly, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears. Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant. While both advantages create a strong barrier to entry, we investigate whether either one is a strong barrier alone, and which one is stronger. Our findings provide a quantitative insight into the classic ``first mover advantage" phenomenon in the digital economy.

@@

%\begin{table}[ht]
%\centering
%\begin{tabular}{cc|c|c|c|}
%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
%            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} &
%            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
%Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} &
%\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
%            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} &
%            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
%\end{tabular}
%\caption{Information Erased Experiment Heavy Tail}
%\label{info_erase}
%\end{table}
%
%\begin{table}[ht]
%\centering
%\begin{tabular}{cc|c|c|c|}
%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
%            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} &
%            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
%Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} &
%\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
%            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} &
%           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
%\end{tabular}
%\caption{Reputation Erased Experiment Heavy Tail}
%\label{rep_erase}
%\end{table}

@@

\begin{finding}\label{barrier-find}
\textit{
(a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances. (b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS. (c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
}
\end{finding}
\OMIT{ %%%%%
\begin{finding}\label{barrier-find}
The incumbent with either data or reputational advantage alone gets a substantially larger share of agents in the competition period, compared to the said share under permanent duopoly. This holds across all algorithms and all MAB instances.  When the incumbent chooses $\TS$, the data advantage serves as a larger barrier to entry compared to the reputation advantage, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}
} %%%%%
% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

Our intuition for Finding~\ref{barrier-find}(b) is as follows. Suppose the incumbent switches from \DG to \TS. This switch allows the incumbent to explore actions more efficiently -- collect better data in the same number of rounds -- and therefore should benefit the data advantage. However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage. One take-away is that the data advantage is not just about data quantity but also data quality.

\OMIT{ %%%%%
\textit{that may be very costly to explore in competition}.
In contrast, purposeful exploration in \TS may damage reputation in the short run. in competition may have immediate adverse reputational consequences. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.
} %%%%%%

\OMIT{ %%%%%
Why is it that data and reputation alone provide such large barriers to entry? Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation. Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences. It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
} %%%%%%

\OMIT{ %%%%%
\gadelete{
The reason \gadelete{for the finding} \gaedit{that we see that the data advantage is greater
under $\TS$} is that the early entry allows the incumbent to do
exploration without immediate reputational consequence in the early
rounds and so, even when reputation is erased, the incumbent has
already acquired sufficient information that simply exploiting that
information allows her to quickly recover her
reputation. \swcomment{which part are you explaining?} However, while
the reputational advantage gives the incumbent some cushion for doing
exploration the incumbent is still vulnerable to getting unlucky while
doing exploration and suffering the reputational consequences. Under all of the incumbent algorithms,
the incumbent obtains a data advantage by the mere fact that it is around in the market and gets data from consumers. However, we see that the data advantage is stronger under \TS since \TS acquires data that is hard to acquire under competition.
We only observe this under $\TS$ since
\TS not only acquiresThis
only happens under $\TS$ since $\TS$, by construction, is more
intelligent about the data that it acquires compared to $\DEG$ and
$\DG$.\swcomment{more intelligent about the data means?}
}
} %%%%%%%

\OMIT{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained. We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.}

>>>>

\section{Data as a Barrier to Entry}\label{section:6}

\gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share. In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}

Under temporary monopoly, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears. Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant. \gaedit{First, our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy. Second, our findings provide a viewpoint in terms of thinking about the role that data can play as a competitive advantage.}

@@

@@

\begin{finding}\label{barrier-find}
\textit{\\
(a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances. \\(b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS. \\(c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
}
\end{finding}

Our intuition for Finding~\ref{barrier-find}(b) is as follows. Suppose the incumbent switches from \DG to \TS. This switch allows the incumbent to explore actions more efficiently -- collect better data in the same number of rounds -- and therefore should benefit the data advantage. However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.

\gaedit{How does this help us in thinking about data as a barrier to entry? The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data. In particular, in the incumbency period \TS gathers data that may be hard for the entrant to gather during competition due to its potential reputational consequences (i.e. engaging in exploration), but which are needed in order to learn. If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker. This insight is what differentiates our results from the static viewpoint of \cite{varian2018artificial, bajari2018impact}. While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad. Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier. This effect only shows itself when thinking about the issue from the perspective of competition dynamics and endogenous data acquisition.}

### 2018-10-18 paper/content/introduction.tex
aaai -> www
somewhat major revision

Platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty. While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, even putting aside the deployment/maintenance costs. In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term. This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.

We ask whether competition incentivizes adoption of ``better" algorithms for exploration. We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its market share (the fraction of users choosing this firm).

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}
      \draw[->] (-.5,0) -- (7,0) node[right] {};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
      \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
     \node[below] at (0.7, -0.2) {\footnotesize monopoly};
     \node[below] at (3, 0) {\footnotesize early \\ entry};
     \node[below] at (4.5, -0.2) {\footnotesize duopoly};
     \node[below] at (6, 0) {\footnotesize late \\entry};
 \end{tikzpicture}
 \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
\label{fig:inverted-U}
\end{center}
\end{figure}

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant. Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``early entry" (in the temporary monopoly) to permanent duopoly to ``late entry" (temporary monopoly from the incumbent's perspective). We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry". This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run. The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).

Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}). This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. One take-away is that the data advantage is not just about data quantity but also data quality.

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.

>>>>

Platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty. While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, even putting aside the deployment/maintenance costs. In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term. This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth. \gaedit{This ``data feedback" loop has been discussed in policy circles and in the business literature, but, to our knowledge, our paper is the first to formally provide a model that allows us to understand this effect.}

\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users \gaedit{on a single dimension, quality}, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}
      \draw[->] (-.5,0) -- (7,0) node[right] {};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
      \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
     \node[below] at (0.7, -0.22) {\footnotesize monopoly};
     \node[below] at (3, -0.2) {\footnotesize incumbent};
     \node[below] at (4.5, -0.22) {\footnotesize duopoly};
     \node[below] at (6, -0.2) {\footnotesize entrant};
 \end{tikzpicture}
 \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
\label{fig:inverted-U}
\end{center}
\end{figure}

We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant. Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (first-mover in duopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late arriver in duopoly)}. \gaedit{We find that, even in a duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late entrant in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run. \gaedit{We find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}

Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}). This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}. \gaedit{The mechanism that leads to this behavior in our model is different from the existing literature in economics since the disincentives to ``innovate" arise entirely from the reputational consequences of exploration in competition and not from R\&D costs (which are absent from our model).}

\gaedit{Our findings show that the data feedback loop arises either when a firm deploys a ``naive" greedy learning algorithm (under simultaneous entry) or has an initial temporary monopoly in the market. The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").

Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest. While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation. We argue that, to reason about this question, one needs to consider what the impact of better data has in terms of competition dynamics and when data collection is endogenous and has consequences, especially when firms need to learn while competing. Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters. In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}

\OMIT{Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. One take-away is that the data advantage is not just about data quantity but also data quality.}

@@

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
\gaedit{There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}). The latter is not about competing
bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".

The study of whether data can give a significant competitive advantage has been discussed theoretically in \cite{varian2018artificial, lambrecht2015can} and empirically in \cite{bajari2018impact}. These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role, whereas we ask how additional data helps while firms simultaneously solve a machine learning problem and competing for a market of consumers. In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm. Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}. }

### 2018-10-18 paper/content/inverted_u.tex
somewhat major revision

Our main experiments are with the duopoly game defined in Section~\ref{sec:model}. As the ``intensity of competition" varies from permanent monopoly to ``early entry" to permanent duopoly to ``late entry", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}. More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS. We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances. For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}). Note that \DG is trivially the dominant strategy under permanent monopoly.

@@

\subsubsection{Inverted-U relationship.}
We interpret our findings through the lens of the inverted-U relationship between the ``intensity of competition" and the ``quality of technology". The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking. The highest levels are permanent duopoly and ``late entry" (temporary monopoly from the entrant's perspective). We see that \DG is incentivized for exploration-disadvantaged MAB instances. In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``late entry".%
\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one. For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
Finally, the middle level of competition, ``early entry" (temporary monopoly for the entrant) creates strong incentives for \TS. In  stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.

@@

Our intuition for why ``early entry" creates more incentives for exploration is as follows. During the temporary monopoly period, reputation costs of exploration vanish. Instead, the firm wants to improve its performance as much as possible by the time competition starts. Essentially, the firm only faces a classical explore-exploit tradeoff, and is incentivized to choose algorithms that are best at optimizing this tradeoff.

@@

>>>>

Our main experiments are with the duopoly game defined in Section~\ref{sec:model}. As the ``intensity of competition" varies from permanent monopoly to ``incumbent" to permanent duopoly to ``entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}. More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS. We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances. For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}). Note that \DG is trivially the dominant strategy under permanent monopoly.

@@

\subsubsection{Inverted-U relationship.}
We interpret our findings through the lens of the inverted-U relationship between the ``intensity of competition" and the ``quality of technology". The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking. The highest levels are permanent duopoly and ``entrant" (temporary monopoly from the entrant's perspective). We see that \DG is incentivized for exploration-disadvantaged MAB instances. In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``entrant".%
\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one. For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
Finally, the middle level of competition, ``incumbent" (\gaedit{duopoly with an initial temporary monopoly}) creates strong incentives for \TS. In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.

@@

Our intuition for why incumbency creates more incentives for exploration is as follows. During the temporary monopoly period, reputation costs of exploration vanish. Instead, the firm wants to improve its performance as much as possible by the time competition starts. Essentially, the firm only faces a classical explore-exploit tradeoff, and is incentivized to choose algorithms that are best at optimizing this tradeoff.

@@

\gaedit{
\subsubsection{Welfare}

What do our results mean for the welfare of consumers? In our model we define consumer welfare in terms of pseudo-regret. Following the bandits literature we define pseudo-regret as being the difference between the true mean of the best action and the true mean of the action that is selected. We equate consumer welfare with \textit{market} pseudo-regret where market pseudo-regret is the sum of the pseudo-regret of the actions that are experienced in the competition game\footnote{Note that smaller pseudo-regret is better for consumers}.

\begin{figure}
\includegraphics[scale=0.3]{figures/unif_eq_welfare}
\includegraphics[scale=0.3]{figures/ht_eq_welfare}
\caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures. Note that welfare at $t = 0$ incorporates the regret incurred during the incumbent and warm start periods.}
\label{eq_regret}
\end{figure}

\begin{finding}
\textit{ \\ (a) Consumer welfare is highest under temporary monopoly. \\
(b) Increasing the number of firms in the market playing \DG weakly decreases consumer welfare.
}
\end{finding}

Figure \ref{eq_regret} displays the simulated welfare when the equilibrium strategies are played. Consumer welfare is \textit{better off} in the temporary monopoly case than in the duopoly case. However, this was expected from the fact that the incumbent is incentivized to play \TS. The welfare in the temporary monopoly case is close to the welfare to having a single firm playing \TS.

What is surprising is that welfare under monopoly and duopoly are very similar, with monopoly being marginally better than duopoly. Recall that in equilibrium in these market structures all firms play \DG. One would think that welfare would be weakly increasing with the number of firms playing \DG since, from the perspective of the market, if adding a firm playing \DG results in that firm, by chance, identifying a better arm than the existing firms then that firm ought to attract more consumers and increase market welfare. If the firm identifies a worse arm than the rest then one ought to expect consumers to not select this firm and market welfare would be unchanged. Indeed, adding sufficiently many firms one would expect that eventually one firm by chance would discover the best arm quickly and the market welfare could even beat that of a stand-alone firm playing \TS.

\begin{figure}
\includegraphics[scale=0.3]{figures/eeog_vs_num_ht}
\includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
\caption{Average welfare and EEOG as we increase the number of firms playing \DG}
\label{many_firm_welfare}
\end{figure}

To test if it is possible that increasing competition past a duopoly leads to better welfare than under temporary monopoly we run simulations where we increase the number of firms playing \DG. Figure \ref{many_firm_welfare} reports the average welfare across these simulations. Welfare not only does not get better, \textit{but is even worse} as we increase the number of firms. Correspondingly, we track the average \Eeog in each of the simulations and notice that it is \textit{increasing} as the number of firms increases. As we increase the number of firms, a single individual firm gets fewer consumers for learning as consumers switch between firms more often. Additionally, bad luck with rewards or arm selection is punished even more harshly as we increase the number of firms. This further strengthens the intuition from our previous result, which is that learning may be hard under competition since accidental exploration and mistakes are punished harshly but are necessary for learning.}

### 2018-10-18 paper/content/model.tex
somewhat major revision

\subsubsection{Agents.}
Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward, without attempting to influence the firms' learning algorithms or rewards of the future users. Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.

Concretely, each firm has a \emph{reputation score}, and each agent's choice is driven by these two numbers. We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly). The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.

@@

For concreteness, we fix one algorithm from each class. Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen. For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge). We consider:

>>>>

\subsubsection{Agents.}
\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the reward experienced by the consumer\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}. Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward \gaedit{(i.e. select the firm with the highest quality)}, without attempting to influence the firms' learning algorithms or rewards of the future users. Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.

Concretely, each firm has a \emph{reputation score}, and each agent's choice is driven by these two numbers. We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly). The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm\gaedit{\footnote{One can think of this as similar to ratings aggregations on many online platforms such as Yelp, Tripadvisor, etc.}}.

@@

For concreteness, we fix one algorithm from each class. Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen. \gaedit{We make this restriction since we are mainly interested in if there is a qualitative differences between what we know is good in isolation and what is incentivized in competition.} For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge). We consider:
